An efficient acyclic contact planner for multiped robots

### 2016-12-09 body/intro.tex
ongoing intro

Planning the locomotion of a multipedal robot in a complex environment - made of stairs, constraining obstacles or uneven surfaces - is still an open issue. Most multipedal locomotion systems focus on cyclic walking gaits \citep{Kajita03a}, but executing this behavior on constraining environments is dangerous if not impossible. In such a context, the robot has to perform a particular class of motion planning called acyclic contact planning, in which the robot must be in contact with the environment at every configuration to maintain the static equilibrium.

The lack of solution to this planning problem has been highlighted by Atkeson \citeauthor{atkensondarpa}. Actually, the current approach to locomotion planning tends to avoid obstacles as much as possible, instead of using them by making contacts to facilitate motion. One main reason is that contact planning is algorithmically difficult, both because of the proximity to obstacles which makes tedious the sampling of collision-free configurations, and because of the foliation of the configuration space where zero-measure manifolds intersect in a combinatorial manner \citep{simeon-manipulation-04}.

Previous contributions that embrace the combinatorial provide complete approaches, not applicable in practice because they require hours of computations \citep{Bretl:2006:MPM:1124573.1124585}.
More recent successes use a local solution, resulting in more reasonable computation times (still far from real time), at the cost of dynamically inconsistent behaviors \citep{Mordatch:2012:DCB:2185520.2185539}.
Neither global nor local methods present satisfying performances, because planning simultaneously the robot trajectory and the contacts that allow
its execution is a very hard problem.

As suggested by \cite{Bouyarmane2009}, we believe that these two problems can be treated sequentially within a global planner.
We go further and claim that this can be done at a much smaller cost, provided we can formulate a computationally-efficient condition for \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} of the root configuration (i.e. there exists a joint configuration such that the robot is in static equilibrium).
This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
This approximation results from a trade-off between a necessary and a sufficient condition for \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

In the remainder of this introduction, we discuss further the current state of the art. This allows us to situate more precisely our contributions.

@@

The issue of planning acyclic contacts was first completely described by Bretl et al. in their seminal paper  \citep{Bretl:2006:MPM:1124573.1124585}. The issue requires the simultaneous handling of two problems, $\mathcal{P}_1$: planning a relevant guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of acyclic equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, not addressed in this work, then consists in interpolating a complete motion between two postures of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in particular situations, typically limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

Most of the papers that followed the work of \citeauthor{Bretl:2006:MPM:1124573.1124585} have explored alternative formulations to handle the combinatorial issue. Two main directions have been explored. \textbf{On one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence. A complete example of the potential offered by such approaches was proposed \citep{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To keep reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the computation time is far from real-time  (about 1 minute of computation for a sequence of 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. \citeauthor{DBLP:conf/humanoids/DeitsT14} propose to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered. 
\citeauthor{dai2014whole} extend the work of \citeauthor{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for more complex interactions.
In addition to the practical limits of the current implementations, a major drawback of these optimization-based approaches is that they only offer local convergence when applied to acyclic contact planning.

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling is shown by \citeauthor{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$). \Pb is addressed as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution can then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. \citeauthor{Bouyarmane2009} precisely focus on automatically computing a guide path with guarantees of \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}, by extending key frames of the path into whole-body configurations in static equilibrium. Randomly-sampled configurations are projected into the contact sub-manifold using a inverse-kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key positions in the path. \citeauthor{7140082} also propose a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with 
a discretized environment. This planning phase does not account for collisions, implying that replanning is required in case of failure. This approach is efficient 
in the demonstrated scenarios. In highly-constraining cases such 
as the car egress scenario that we address here, we believe that including collision constraints in the planning is a requirement.

@@

\noindent \textbf{Regarding $\mathcal{P}_1$}, we define and use the \textit{reachability condition} to plan the guide path. The \textit{reachability condition}, a computationally-efficient approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}. This step is fast because the planning happens
in a low-dimensional space, and does not require explicit contact computations.

@@

As a result, we propose contributions to both problems \Pa and $\mathcal{P}_2$.
\begin{itemize}
\item A low-dimensional, efficient sampling-based approach to plan guide paths.
\item A very efficient and general implementation of an acyclic contact planner, the first one compatible with \gls{interactive} applications.
\item Four heuristics for contact generation. They bias the planner towards configurations that are in robust static equilibrium, or 
more efficient with respect to the task.
\end{itemize}

>>>>

%~ Planning legged robot locomotion remains an open issue.
As humans do, a legged robot moves by sequentially creating contacts with the environment while avoiding collisions. Finding such a contact sequence is an instance of the motion planning problem called acyclic contact planning.

Once the contact plan is known, efficient approaches can execute the plan and generate a dynamically feasible motion~\citep{Carpentier2016}. However in general
the contact plan is manually designed after long trials and error. Automatically addressing the acyclic contact planning problem is thus a pre-requisite to the deployment of robots in human-centered environments, which are cluttered and comprise constraining obstacles such as stairs.

To make a legged robot walk on a flat ground, efficient approaches based on the capture point~\citep{Pratt2006} can be used to choose the next contact location in a deterministic manner. Choosing the effector to create a contact with is trivial because walk follows a cyclic pattern.
Unfortunately, in the general case planning complex contact interactions is extremely challenging because the determinism disappears.
At any given time a contact choice must be made between infinitely many possibilities (often a combinatorial discrete choice for the effector and contact surface, and a continuous choice for the contact location). Furthermore, the contact manifold is foliated~\citep{simeon-manipulation-04}: a contact choice constrains kinematically and dynamically the possible motions, and there is no analytical way to verify whether this decision brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles. Lastly, the contact manifold has a null measure and is thus impossible to randomly sample. This restriction prevents the use of efficient sampling based planners to solve the problem.

For this reason previous contributions having demonstrated acyclic contact locomotion on a real robot are too computationally expensive~\citep{Bretl:2006:MPM:1124573.1124585}. Aa a result at the DARPA challenge, the robots fell back to unsafe walking patterns to climb stairs, instead of using the provided handrails to faciliate the motion~\citep{atkensondarpa}.

Reaching reasonable computation times is critical for a contact planner to be of any practical use. Our work thus aims at breaking the complexity of the contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the non-determinism of the contact selection problem. First we introduce a low dimensional space, called the \textit{contact-feasible} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact-feasible} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling between the planning of the path and the generation of contacts presents drawbacks and advantages discussed in previous related literature.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

%~ In the remainder of this introduction, we discuss further the current state of the art. This allows us to situate more precisely our contributions.

@@

The issue of planning acyclic contacts was first completely described by Bretl et al. in their seminal paper  \citep{Bretl:2006:MPM:1124573.1124585}. The issue requires the simultaneous handling of two problems, $\mathcal{P}_1$: planning a relevant guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of acyclic equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, not addressed in this work, then consists in interpolating a complete motion between two postures of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in specific situations, limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

Most of the papers that followed the work of \citeauthor{Bretl:2006:MPM:1124573.1124585} have explored alternative formulations to handle the combinatorial issue. Two main directions have been explored. \textbf{On one hand, local optimization of both the root trajectory \Pa and the contact positions $\mathcal{P}_2$} has been used, to trade the combinatorial of the complete problem for a differential complexity, at the cost of local convergence. A complete example of the potential offered by such approaches was proposed by \citeauthor{Mordatch:2012:DCB:2185520.2185539} and successfully applied to a real robot \citep{mordatch2015}. To keep reasonable computation times, the method uses a simplified dynamic model for the avatar. Still, the computation time is far from real-time  (about 1 minute of computation for a sequence of 20 contacts).  A similar approach has been considered for manipulation by \cite{gabicciniisrr15}. \citeauthor{DBLP:conf/humanoids/DeitsT14} propose to solve contact planning globally as a mixed-integer problem, but only cyclic, bipedal locomotion is considered. 
\citeauthor{dai2014whole} extend the work of \citeauthor{Posa:2014:DMT:2568343.2568352} to discover the contact sequence for landing motions, but need to specify
the contacts manually for more complex interactions.
In addition to the limits of the current implementations, a major drawback of optimization-based approaches is that they only offer local convergence.

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling is shown by \citeauthor{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$). \Pb is addressed as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution can then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. \citeauthor{Bouyarmane2009} precisely focus on automatically computing a guide path with guarantees of \textit{equilibrium feasibility}, by extending key frames of the path into whole-body configurations in static equilibrium. Randomly-sampled configurations are projected into the contact sub-manifold using a inverse-kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key positions in the path. \citeauthor{7140082} also propose a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with 
a discretized environment. This planning phase does not account for collisions, implying that replanning is required in case of failure. This approach is efficient 
in the demonstrated scenarios. In highly-constraining cases such 
as the car egress scenario that we address here, we believe that including collision constraints in the planning is a requirement.

@@

\noindent \textbf{Regarding $\mathcal{P}_1$}, we define and use the \textit{reachability condition} to plan the guide path. The \textit{reachability condition} defines a low dimensional space that can be mapped to the contact manifoldcomputationally-efficient approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}. This step is fast because the planning happens
in a low-dimensional space, and does not require explicit contact computations.

@@

As a result, we propose contributions to both problems \Pa and $\mathcal{P}_2$.
\begin{itemize}
\item A low-dimensional, efficient sampling-based approach to plan guide paths.
\item A very efficient and general implementation of an acyclic contact planner, the first one compatible with \gls{interactive} applications.
%~ \item Four heuristics for contact generation. They bias the planner towards configurations that are in robust static equilibrium, or 
%~ more efficient with respect to the task.
\end{itemize}
